Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions

### 2024-08-11 latex/vanishing.tex
almost finished with pazman

\subsubsection{Clusterization in our generic model}
In our generic model, the limit on the number of measurements arises
as a consequence of Carath\'eodory's Theorem\footnote{Let $X \subseteq
\mathbb{R}^n, X \neq \phi$ and denote $\conv (X)$ the convex hull of
$X$. Let $x \in \conv (X)$. Then $x$ is a convex combination of at
most $n+1$ vectors in $X$.} and the concentration on the first $k$
eigenvectors of $\fwd \prcov \fwd^*$. The discussion below extends
arguments presented in \cite[Chapter 3]{silvey} and \cite[Section
  5.2.3]{pronzatoPazman2013}.

We now ignore all but the first $k$ eigenvectors of $\fwd \prcov
\fwd$, as instructed by Theorem \ref{thm:char}. Thus, we can safely
replace $\hilo$ with the $k$-dimensional subspace spanned by the first
$k$ eigenvectors of $\fwd^*\prcov\fwd$. Since this subspace is
isomorphic to $\mathbb{R}^k$, we do not name it and simply denote it
by $\mathbb{R}^k$. Now, let
\begin{equation*}
  \mathcal{M} := \conv \{\meas \meas^* : \meas\in \mathbb{R}^k, \|\meas\|=1\},
  %\hilo^{(k)}\},
\end{equation*}
where $\conv$ denotes the convex hull of a set. Since $\mathcal{M}$ is
a set of positive-definite matrices, it lives in a
$k(k+1)/2$-dimensional subspace of $\mathbb{R}^{k\times k}$. Now note
that $\frac{1}{m}\opt^*\opt \in \mathcal{M}$. By Carath\'eodory's
Theorem, there exists a convex combination
\begin{equation*}
  \frac{1}{m} \opt^*\opt = \sum_{i=1}^{\frac{k(k+1)}{2} + 1} \alpha_i \meas\meas^*,
\end{equation*}
where $\|\meas\|=1, \sum_{i=1}^I\alpha_i = 1$ and $\alpha_i > 0$.

For example, if $k=2$ --- as is the case for the inverse prblem of the
heat equation --- then an optimal design $\opt$ is a combination of at
most $3$ measurement vectors $\meas$. Thus, if the number of allowed
measurements $m$ is larger than $3$, then there exists an optimal
clustered design.

>>>>

\subsubsection{Clusterization in our generic model}
We can gain further insight to clusterization in our generic model,
from Carath\'eodory's Theorem and the concentration on the first $k$
eigenvectors of $\fwd \prcov \fwd^*$. We present a short discussion
adapting arguments presented in \cite[Chapter
  3]{silveyOptimalDesign1980} and \cite[Section
  5.2.3]{pronzatoPazman2013}.

Consider $\opt$ a D-optimal design under our generic model.  As
instructed by Theorem \ref{thm:char}, we ignore all but the first $k$
eigenvectors of $\fwd \prcov \fwd$. Thus, we replace $\hilo$ with
$\hilo^{(k)}$ --- the $k$-dimensional subspace spanned by the first
$k$ eigenvectors of $\fwd^*\prcov\fwd$. We keep in mind that
$\hilo^{(k)}$ is isomorphic to $\mathbb{R}^k$ and let
\begin{equation*}
  \mathcal{M} := \conv \{\meas \meas^* : \meas\in \hilo^{(k)}, \|\meas\|=1\},
  %\hilo^{(k)}\},
\end{equation*}
where $\conv$ denotes the convex hull of a set. The set $\mathcal{M}$
contains only positive-definite matrices, hence it lives in a
$k(k+1)/2$-dimensional vector space. Since $\opt^*\opt = \sum_{i=1}^m
\meas_i\meas_i^*$ for $\meas_i \in \hilo^{(k)}$, it is easy to verify
that $\frac1m \opt^*\opt \in \mathcal{M}$.  Now, recall
Carath\'eodory's Theorem:
\begin{theorem}[Carath\'eodory]
  Let $X \subseteq \mathbb{R}^n, X \neq \phi$ and denote $\conv (X)$
  the convex hull of $X$. For every $x \in \conv (X)$, $x$ is a convex
  combination of at most $n+1$ vectors in $X$.
\end{theorem}
Carath\'eodory's Theorem implies that there exist $\meas_i$ and
$\alpha_i$ such that
\begin{equation*}
  \opt^*\opt = \sum_{i=1}^I \alpha_i \meas_i\meas_i^*,
\end{equation*}
where $\|\meas\|=1, \sum\alpha_i = m, \alpha_i \geq 0$ and $I =
\frac{k(k+1)}{2} + 1$. We can thus write $\opt$ as (note that we do
not require $\meas_i$ to be orthogonal to each other):

\[
\opt =
\left[
  \begin{array}{ccc}
    \horzbar & \sqrt{\alpha_i} \meas^*_1 & \horzbar \\
    \horzbar & \sqrt{\alpha_2} \meas_2^* & \horzbar \\
             & \vdots    &          \\
    \horzbar & \sqrt{\alpha_I} \meas_I^* & \horzbar \\
  \end{array}
\right]
\]

Therefore, if $m > \frac{k(k+1)}{2} + 1$, then at least one
coefficient $\alpha_i > 1$. Note that the abovementioned $\opt$ is not
a valid design (since its rows do not have unit norm,
i.e.~$\|\sqrt{\alpha_i}\meas\| \neq 1$). However, such representation
provides another, deep view of clusterization.

%% For example, if $k=2$, there exists an optimal design $\opt$ that can
%% be represented be as a linear combination of at most $(2+1)/2 + 1 = 4$
%% measurement vectors $\meas$, with nonnegative coefficients whose
%% squares sum to $m$. If, the number of allowed measurements $m=5$, we
%% can write $\opt = \sum_{i=1}^4 \alpha_i \meas_i$, where $\alpha_i > 1$
%% for some $i\in \{1,2,3,4\}$. Since we put weight $>1$ on the
%% $i\textsuperscript{th}$ measurement, we can view such a design as
%% clustered.
